Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-08-04 Paper/assa.tex
..

\subsection{Active subspaces}
\label{sub:ac}
As mentioned in the introduction, the active subspace is low-dimensional subspace
that constitutes important directions in a model's input
space~\cite{Constantine:2015}. The effective variability in the model output, $f$
due to uncertain inputs, $\bm{\theta}$ in the physical space, is predominantly captured
along these directions. Each component of $\bm{\theta}$
can be mapped to a corresponding component of $\bm{\xi}$ in $\Omega\in\mathbb{R}^{N_p}$
($N_p$: number of uncertain inputs)
using a linear transformation.  Hence, from this point onwards, we denote $f$ as $f(\bm{\xi})$.
Mathematically, for a given $f$, and a set of inputs, $\bm{\xi}$,
these directions are the eigenvectors of a positive
semi-definite matrix, $\mat{C}$, defined as follows: 
%
\be
\mat{C} = \int_\Omega (\nabla_{\bm{\xi}}f)(\nabla_{\bm{\xi}}f)^\top dP_{\bm{\xi}}
\label{eq:C}
\ee
%
where 
$dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
distribution of $\bm{\xi}$. Note that the function, $f$ is considered to be continuous 
and differentiable in the physical space, $\bm{\theta}$. %Hence, it is possible that a given $f(\bm{\xi})$ might not admit an active
%subspace. However, it is of remarkable interest to investigate if one exists
%since the subspace could be exploited to reduce the dimensionality of the
%problem and hence the associated computational effort.
Here 
$\nabla_{\bm{\xi}}f$ denotes the gradient vector. 
%with individual components
%being partial derivatives of $f$ with respect to the $i^\text{th}$ input, $\xi_i$. 
Since $\mat{C}$ is symmetric and
positive semi-definite, it admits a spectral decomposition:
%
\be
\mat{C} = \mat{W}\mat{\Lambda}\mat{W}^\top
\ee
%
where $\mat{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with
$\lambda_i$'s sorted in descending order
\[
     \lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_\Np \geq 0.
\] 
The eigenpairs are partitioned about the $i^{\text{th}}$ eigenvalue such that
the ratio, $\left(\frac{\lambda_i}{\lambda_{i+1}}\right)$ $\gg$ 1 as follows:
\be
 \mat{W} = [\mat{W}_1~\mat{W}_2],~~\mat{\Lambda} = \begin{bmatrix}\mat{\Lambda}_1 & \\  &
  \mat{\Lambda}_2 \\
\end{bmatrix}
\ee
 %
 where $\mat{W}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as
the active subspace, and $\mat{\Lambda}_1$ is the corresponding set of
eigenvalues.

>>>>

\subsection{Active subspaces}
\label{sub:ac}
As mentioned in the introduction, the active subspace is low-dimensional subspace
that constitutes important directions in a model's input
space~\cite{Constantine:2015}. The effective variability in the model output, $f$
due to uncertain inputs, $\bm{\theta}$ in the physical space, is predominantly captured
along these directions. Each component of $\bm{\theta}$
is mapped to a corresponding value in $\Omega\in\mathbb{R}^{N_p}$
($N_p$: number of uncertain inputs)
using a linear transformation.  The set of projected values in $\Omega$ is denoted as $\bm{\xi}$.
Hence, from this point onwards, we denote $f$ as $f(\bm{\xi})$.
Mathematically, for a given $f$, and a set of inputs, $\bm{\xi}$,
these directions are the eigenvectors of a positive
semi-definite matrix, $\mat{C}$, defined as follows: 
%
\be
\mat{C} = \int_\Omega (\nabla_{\bm{\xi}}f)(\nabla_{\bm{\xi}}f)^\top dP_{\bm{\xi}}
\label{eq:C}
\ee
%
where 
$dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
distribution of $\bm{\xi}$. Note that the function, $f$ is considered to be continuous 
and differentiable in the physical space, $\bm{\theta}$. %Hence, it is possible that a given $f(\bm{\xi})$ might not admit an active
%subspace. However, it is of remarkable interest to investigate if one exists
%since the subspace could be exploited to reduce the dimensionality of the
%problem and hence the associated computational effort.
Here 
$\nabla_{\bm{\xi}}f$ denotes the gradient vector. 
%with individual components
%being partial derivatives of $f$ with respect to the $i^\text{th}$ input, $\xi_i$. 
Since $\mat{C}$ is symmetric and
positive semi-definite, it admits a spectral decomposition:
%
\be
\mat{C} = \mat{W}\mat{\Lambda}\mat{W}^\top
\ee
%
where $\mat{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with
$\lambda_i$'s sorted in descending order
\[
     \lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_\Np \geq 0.
\] 
The eigenpairs are partitioned about the $i^{\text{th}}$ eigenvalue such that
the ratio, $\left(\frac{\lambda_i}{\lambda_{i+1}}\right)$ $\gg$ 1 as follows:
\be
 \mat{W} = [\mat{W}_1~\mat{W}_2],~~\mat{\Lambda} = \begin{bmatrix}\mat{\Lambda}_1 & \\  &
  \mat{\Lambda}_2 \\
\end{bmatrix}
\ee
 %
 where $\mat{W}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as
the active subspace, and $\mat{\Lambda}_1$ is the corresponding set of
eigenvalues.
